Hydras, Ordinals \& Co.\\ A library in Coq of entertaining formal mathematics

### 2021-10-29 doc/chapter-powers.tex
Add symmetric notations for hydras/gaia. Fix a few typos
more snippets in gaia-chapter
simplifying latex macros

In~\cite{BC04}, several exercises~\footnote{Exercises 9.8 (page 270), 9.10 (page 271), 9.15 (page 276), 9.17 (page 284), and 15.8 (page 418).}
present ways to compute Fibonacci numbers, with the less number of recursive calls  as possible. Please note that these optimizations and the formal proof of their correctness are \emph{ad-hoc}, \emph{i.e.}, exclusively written for the
Fibonacci numbers.
In contrast, the optimizations we present in this document apply, in their vast majority, \emph{generic} techniques of efficient computation of powers in a monoid. 
This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}. This original version is available on
\href{../theories/html/additions.fib.html}{additions.fib}.

@@

The following tests show various examples of chains for the same exponent, using different strategies. The dichotomic strategy seems clearly to be the winner (at least on this sample)\footnote{For efficiency's sake, we commented out some (very) long computations. You may uncomment them freely in your own copy. For the same reason, we put a verbatim trace instead of \textit{Alectryon} output}.

>>>>

In~\cite{BC04}, several exercises~\footnote{Exercises 9.8 (page 270), 9.10 (page 271), 9.15 (page 276), 9.17 (page 284), and 15.8 (page 418).}
present ways to compute Fibonacci numbers, with the less number of recursive calls  as possible. Please note that these optimizations and the formal proof of their correctness are \emph{ad-hoc}, \emph{i.e.}, exclusively written for the
Fibonacci numbers.
In contrast, the optimizations we present in this document apply, in their vast majority, \emph{generic} techniques of efficient computation of powers in a monoid. 
This example of Fibonacci numbers has been developed with Yves Bertot, who wrote a first version with \texttt{SSreflect/Mathcomp}~\cite{MCB}. This original version is available on hydra-battles in the module \href{../theories/html/additions.fib.html}{additions.fib}.

@@

The following tests show various examples of chains for the same exponent, using different strategies. The dichotomic strategy seems clearly to be the winner (at least on this sample)\footnote{For efficiency's sake, we commented out some (very) long computations. You may uncomment them freely in your own copy. For the same reason, we put a verbatim trace instead of \alectr  output}.

### 2021-10-29 doc/chapter-primrec.tex
fix typos
simplifying latex macros

\begin{remark}
In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}. Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement. Our example of page\pageref{sect:bigfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
\texttt{Defined}s.

@@

Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an IDE or software like Alectryon.
%%
%%% To move the path chapter
%%%%

>>>>

\begin{remark}
In the library \texttt{primRec}, all these lemmas are opaque (registered with \texttt{Qed}. Thus they do not allow the user to look at the witness of a proof of a \texttt{isPR} statement. Our example of page~\pageref{sect:bigfac} was built using a  copy of \texttt{primRec.v} where many \texttt{Qed}s have been replaced with
\texttt{Defined}s.

@@

Our proof of this lemma is not trivial at all, it uses some properties of the Ketonen-Solovay's toolkit. We advise the reader to explore this proof, with the help of an IDE or software like \alectr.
%%
%%% To move the path chapter
%%%%

### 2021-10-29 doc/gaia-chapter.tex
simplifying latex macros

This proof, as it appears in \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, is quite dense and qualified as ``a bit tricky''. We present here its main steps with the help of the alectryon tool.

@@

\item
  The term is equal to
  \texttt{(cons a n'' b'')}, where $n''<n$ and $b''<\omega^a$.
  This term is less than the accessible \texttt{(cons a n b'')}.

@@

 Then we can apply the hypothesis \texttt{He}.

@@

In every case, the considered ordinal is accessible.

>>>>

This proof, as it appears inw, is quite dense and qualified as ``a bit tricky''. We present here its main steps with the help of the \alectr tool.

@@

\item
  The term is equal to
  \texttt{(cons a n'' b'')}, where $n''<n$ and $b''<\omega^a$.

@@

  By \texttt{Hd} and \texttt{He}, and a few technical lemmas about normal forms,   the term  \texttt{(cons a n b'')} is accessible, and so is \texttt{(cons a n'' b'')}.

@@

In every case, the considered ordinal is accessible. By the definition of accessibility, the ordinal \texttt{(cons a n b)} is accessible, which ends the proof. The reader is invited to replay the full proof in  \href{https://github.com/coq-community/gaia/blob/master/theories/ssete9.v}{gaia.ssete9.v}, and to compare with
the proof in Section~\ref{sec:strongly-accessible}.

### 2021-10-29 doc/hydras.tex
and Cie -> and Co. (book title)
and -> ampersand (book)
title change
change title and subtitle
add a new chapter (Gaia)
simplifying latex macros

\documentclass[a4paper]{book}
\usepackage{alltt}
\usepackage{multind}
\usepackage{hyperref}
\usepackage{url}
\usepackage{xcolor,fancyvrb, mdframed}
\usepackage{amsmath}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{float}
\usepackage{threeparttable}
\usepackage{fontspec}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{placeins}
\usepackage{tikz}
\usepackage{tikzsymbols}
\usetikzlibrary{arrows}
\usepackage{graphicx}
\usepackage{amsmath,mathdots,dsfont}
\usepackage{amssymb}
\usepackage{newunicodechar}
\usepackage{theorem}

\usepackage{texments}
%%% for movies by alectryon
\usepackage{movies/snippets/assets/alectryon}
\usepackage{movies/snippets/assets/pygments}

@@

\title{Hydra Battles and Cie}
\makeindex{coq}
\makeindex{maths}
\makeindex{hydras}
\makeindex{additions}
\makeindex{primrec}
\begin{document}
\maketitle

@@

This document is \emph{not} an introductory text for \coq{}, and there are many aspects of this proof assistant that are not covered. 
 The reader should already have some basic experience with the \coq{} system. The Reference Manual and several tutorials are available on the \coq{} website~\cite{Coq}. The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.

@@

\subsection{Documenting theories with Alectryon}

Quotations of \coq{} source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \emph{Alectryon} tool~\cite{alectryonpaper, alectryongithub}.
Many thanks to Jérémy Damour, Clément Pit-Claudel  and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.

@@

\subsubsection{Using Alectryon}

@@

\include{part-hydras}
\include{chapter-primrec}
\part{A few  certified algorithms}

>>>>

\documentclass[twoside,a4paper]{book}
\usepackage{alltt}
\usepackage{multind}
\usepackage{hyperref}
\usepackage{url}
\usepackage{xcolor, fancyvrb, mdframed}
\usepackage{amsmath}
%\usepackage{fancyhdr}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{float}
\usepackage{threeparttable}
\usepackage{fontspec}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{placeins}
\usepackage{tikz}
\usepackage{tikzsymbols}
\usetikzlibrary{arrows}
\usepackage{graphicx}
\usepackage{amsmath,mathdots,dsfont}
\usepackage{amssymb}
\usepackage{newunicodechar}
\usepackage{theorem}
\usepackage{xspace}
\usepackage{texments}
%%% for movies by alectryon
\usepackage{movies/snippets/assets/alectryon}
\usepackage{movies/snippets/assets/pygments}

% Prevent breaks in the middle of syntactic units
\let\OldPY\PY
\def\PY#1#2{\OldPY{#1}{\mbox{#2}}}

@@

\title{Hydras, Ordinals \& Co.\\ A library in Coq of entertaining formal mathematics}
%\pagestyle{fancy}
%\fancyhead{Hydras, Ordinals, \& Co.}
\makeindex{coq}
\makeindex{maths}
\makeindex{hydras}
\makeindex{additions}
\makeindex{primrec}
\begin{document}
\maketitle

@@

This document is \emph{not} an introductory text for \coq, and there are many aspects of this proof assistant that are not covered. 
 The reader should already have some basic experience with the \coq system. The Reference Manual and several tutorials are available on the \coq website~\cite{Coq}. The first chapters of textbooks like \emph{Interactive Theorem Proving and Program Development}~\cite{BC04}, \emph{Software Foundations}~\cite{SF} or  \emph{Certified Programming with Dependent Types} ~\cite{chlipalacpdt2011} will give you the right background.

@@

\subsection{Documenting theories with \alectr}

Quotations of \coq source and answers are progressively replaced from copy-pasted \emph{verbatim} to automatically generated \emph{LaTeX} blocks, using Clément Pit-Claudel's \alectr  tool~\cite{alectryonpaper, alectryongithub}.
Many thanks to Jérémy Damour, Clément Pit-Claudel  and Théo Zimmermann who designed tools for maintaining consistency between the always evolving \coq{} modules and documentation written in \emph{LaTeX}.

@@

\subsubsection{Using \alectr}

@@

\include{part-hydras}
\include{chapter-primrec}
\include{gaia-chapter}
\part{A few  certified algorithms}
